---
title: "Prediction Challange 1"
author: "David Parsons"
date: "March 31, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
options(scipen=999)
```

---

[Rutgers Data 101 Blog](http://data101.cs.rutgers.edu/?q=blog)

[My Data Exploration Homepage](home.html)

[Rutgers Data 101 Homepage](http://data101.cs.rutgers.edu/)

---

## Introduction

This investigation is in response to a course assignment for my CS 142 class. We we're provided with a synthetic data set that had artificial patterns created with in it for us to find.

## The Data

Data Provided by Professor Imenlinski as part of a CS 142 Assignment.

```{r The Data}
train <- read.csv("M2017_train.csv")

print(summary(train))
```
```{r Adding data}
train[,7] <- ""
colnames(train)[7] <- "GRADELEVEL"
train$GRADELEVEL <- as.numeric(0)
train$GRADELEVEL[train$GRADE == "A"] <- as.numeric(5) 
train$GRADELEVEL[train$GRADE == "B"] <- as.numeric(4) 
train$GRADELEVEL[train$GRADE == "C"] <- as.numeric(3) 
train$GRADELEVEL[train$GRADE == "D"] <- as.numeric(2) 
train$GRADELEVEL[train$GRADE == "F"] <- as.numeric(1) 

train[,8] <- ""
colnames(train)[8] <- "SCORELEVEL"
train$SCORELEVEL <- cut(train$SCORE, 10, c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"))

train[,9] <- ""
colnames(train)[9] <- "IDLEVEL"
train$IDLEVEL <- cut(train$STUDENTID, 10, c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"))

vars = colnames(train)

summary(train)

```

Each data point was representative of a single person and how they fit into the following:

    StudentID (unique ID)
    Score (0 - 100)
    Grade (A - F)
    Asks Questions (Always, Never, Sometimes)
    Leaves Early (Always, Never, Rarely)
    Participation (0 - 1)


## Investigation

The following is the step by step thought process I followed in exploring this data set and arriving at the conclusion I did.

### Happiness, Income, & Age

I started by creating a scatter plot of each respondents Happiness against both Income and Age


```{r Simple Plots}

scatPlot <- function(x, y, frm){
  
  title = paste(x,y, sep = " vs ")
  plot(frm[,x], frm[,y], col = rainbow(8), main = title, las = 2, xlab = x, ylab = y)
}

heatPlot <- function(frm, yAxis, xAxis, heat) {
  subCen <- subset(frm)
  print(ggplot(data = subCen, aes(x = subCen[,xAxis], y = subCen[,yAxis]), las = 2)+ labs(x = xAxis, y = yAxis, las = 2) + geom_tile(aes(fill = subCen[,heat]))+
          scale_fill_gradient2(low="darkblue", high="red", guide="colorbar") + theme(axis.text.x = element_text(angle = 90, hjust = 1)))
}

t <- subset(train, train$GRADE == "D")

for(v in vars){
  print(v) 
  barplot(height=table(t[,v]))
}

t <- train

c <- "SCORELEVEL"
for(v in vars){
  if(c != v){
    scatPlot(c, v, t)
  }
}

c <- "SCORELEVEL"
for(v in vars){
  # if(c != v){
    heatPlot(t, c, v, "GRADELEVEL")
  # }
}

```


Age vs Happiness revealed no correlations, Age appears to have no direct effect upon a person's happiness.

But Income vs Happiness shows clear increase in happiness as income increases.

### Income & Happiness "Classes"

Now I wanted to define what Ranges or classes exist within Happiness and Income.

```{r Score Changer}


train[,10] <- ""
colnames(train)[10] <- "FSCORE"
train$FSCORE <- 0

train$FSCORE[train$PARTICIPATION > 0.5] <- (train$FSCORE[train$PARTICIPATION > 0.5]) + 2

train$FSCORE[train$ASKS_QUESTIONS == "always"] <- (train$FSCORE[train$ASKS_QUESTIONS == "always"]) + 1

train$FSCORE[train$LEAVES_EARLY == "never"] <- (train$FSCORE[train$LEAVES_EARLY == "never"]) + 1

train$FSCORE[train$SCORE > 80] <- train$FSCORE[train$SCORE > 80] + 3
train$FSCORE[train$SCORE > 60] <- train$FSCORE[train$SCORE > 60] + 2
train$FSCORE[train$SCORE > 40] <- train$FSCORE[train$SCORE > 40] + 1
train$FSCORE[train$SCORE > 20] <- train$FSCORE[train$SCORE > 20] + 2

summary(train)

train[,11] <- ""
colnames(train)[11] <- "NGRADE"
train$NGRADE <- "F"
train$NGRADE[train$FSCORE > 2] <- "D"
train$NGRADE[train$FSCORE > 4] <- "C"
train$NGRADE[train$FSCORE > 6] <- "B"
train$NGRADE[train$FSCORE > 8] <- "A"

train[,12] <- ""
colnames(train)[12] <- "SAME"
train$SAME <- "False"
train$SAME[train$GRADE == train$NGRADE] <- "True"



```

### Output

```{r Output}
out <- read.csv("M2017_test_students.csv")

out[,6] <- ""
colnames(out)[6] <- "FSCORE"
out$FSCORE <- 0

out$FSCORE[out$PARTICIPATION > 0.5] <- (out$FSCORE[out$PARTICIPATION > 0.5]) + 2

out$FSCORE[out$ASKS_QUESTIONS == "always"] <- (out$FSCORE[out$ASKS_QUESTIONS == "always"]) + 1

out$FSCORE[out$LEAVES_EARLY == "never"] <- (out$FSCORE[out$LEAVES_EARLY == "never"]) + 1

out$FSCORE[out$SCORE > 80] <- out$FSCORE[out$SCORE > 80] + 3
out$FSCORE[out$SCORE > 60] <- out$FSCORE[out$SCORE > 60] + 2
out$FSCORE[out$SCORE > 40] <- out$FSCORE[out$SCORE > 40] + 1
out$FSCORE[out$SCORE > 20] <- out$FSCORE[out$SCORE > 20] + 2


out[,7] <- ""
colnames(out)[7] <- "GRADE"
out$GRADE <- "F"
out$GRADE[out$FSCORE > 2] <- "D"
out$GRADE[out$FSCORE > 4] <- "C"
out$GRADE[out$FSCORE > 6] <- "B"
out$GRADE[out$FSCORE > 8] <- "A"

# TO File

submission<-read.csv("M2017_sample_submission.csv")
submission$GRADE<-out$GRADE
write.csv(submission, 'submission.csv', row.names= FALSE)

```

## Notes and Comments

This attempt went much more smoothly. I've begun to get a better grasp on how to approach analyzing a data set, and have made use of functions and loops to increase my plot speed.

I felt I over thought the last analysis, so I went with a more direct approach this time, and it worked out.

But I think that because this is a synthetic data set, it was easier to over think the analysis. So I look forward to dealing with more random/real world data sets in the future.

Our next project won't involve a synthetic data set, so I look forward to the new challenges that the next step in analysis presents.

---

[Rutgers Data 101 Blog](http://data101.cs.rutgers.edu/?q=blog)

[My Data Exploration Homepage](home.html)

[Rutgers Data 101 Homepage](http://data101.cs.rutgers.edu/)