---
title: "Happiness"
author: "David Parsons"
date: "February 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
options(scipen=999)
```

---

[Rutgers Data 101 Blog](http://data101.cs.rutgers.edu/?q=blog)

[My Data Exploration Homepage](home.html)

[Rutgers Data 101 Homepage](http://data101.cs.rutgers.edu/)

---

## Introduction

This investigation is in response to a course assignment for my CS 142 class. We we're provided with a synthetic data set that had artificial patterns created with in it for us to find.

## The Data

Data Provided by Professor Imenlinski as part of a CS 142 Assignment.

```{r The Data}
hap <- read.csv("HAPPINESS2017.csv")

hap[,8] <- ""
colnames(hap)[8] <- "GEN"
hap$GEN <-  cut(hap$AGE, 5, c("Young", "Adult", "Middle Adult", "Older Adult", "Elderly"))

summary(hap)

vars = colnames(hap)
```

Each data point was representative of a single person and how they fit into the following:

    Idn (unique ID)
    Age
    Country (Spouse Origin)
    Gender
    Immigrant (Binary Immigrant status 1 -> Yes, 0 -> No)
    Income
    Happiness ("Happiness" Level from 1 - 10)


## Investigation

The following is the step by step thought process I followed in exploring this data set and arriving at the conclusion I did.

### Happiness, Income, & Age

I started by creating a scatter plot of each respondents Happiness against both Income and Age


```{r Simple Plots}

scatPlot <- function(x, y, frm){
  
  title = paste(x,y, sep = " vs ")
  plot(frm[,x], frm[,y], col = rainbow(8), main = title, las = 2, xlab = x, ylab = y)
}

scatPlot("AGE", "HAPPINESS", hap)
scatPlot("INCOME", "HAPPINESS", hap)

```


Age vs Happiness revealed no correlations, Age appears to have no direct effect upon a person's happiness.

But Income vs Happiness shows clear increase in happiness as income increases.

### Income & Happiness "Classes"

Now I wanted to define what Ranges or classes exist within Happiness and Income.

I spent more time than I'd like to admit, brute forcing the ranges until the scatter plot was void of those with a Happiness Level equal to their income level.

I knew just from looking at the graph that there were 8 distinct level for both, and rather than trying to use R functions that I didn't fully understand, I just changed the ranges and re-plotted the data adjusting based of the visual data. I should have poured over some documentation and learned how get R to do it for me (the whole point of these projects), but after failing with cut and pretty, I just brute forced it and called it a night. I'll read the R docs when I have to do this in the future.

But in the end I wound up with these new categories and this new subset of data after I removed those with Happiness Levels equal to their Income levels.

```{r Income & Happiness Classes}

hap[,9] <- ""
colnames(hap)[9] <- "INCLEVEL"
classes = c("8th", "7th", "6th", "5th", "4th", "3rd", "2nd", "1st")
hap$INCLEVEL <-  cut(hap$INCOME, breaks = c(0, 36000, 47999, 59999, 72000, 84000, 96000, 108000, 120000), classes)


hap[,10] <- ""
colnames(hap)[10] <- "HAPLEVEL"
classes = c("8th", "7th", "6th", "5th", "4th", "3rd", "2nd", "1st")
hap$HAPLEVEL <-  cut(hap$HAPPINESS, breaks = c(0, 3, 4, 5, 6, 7, 8, 9, 10), classes)

summary(hap)

plotBar <- function(frm, val, cat){
  
  frm.mean <- tapply(frm[,val], frm[,cat], mean)
  barplot(frm.mean, las = 2, main=paste(val, " by ", cat), xlab = cat, ylab = val, col = rainbow(6), las = 2)
}

plotBar(hap, "HAPPINESS", "INCLEVEL")
plotBar(hap, "INCOME", "HAPLEVEL")
plotBar(hap, "HAPPINESS", "INCOME")
```

These bar graphs confirm the clear correlation between an increase in Income and Happiness.

The graph looks almost like a set of stairs. Each step is an increase in Income and it correlates to an increase in "Happiness" and vice versa.

But now I am interested in those outliers, that didn't fit into the steps.

So here I eliminate all respondents that don't have Happiness Level equal to their Income Levels

```{r Outliers}
# Original Set
scatPlot("INCOME", "HAPPINESS", hap)
# Set with the with the same Income & Happiness Levels removed
scatPlot("INCOME", "HAPPINESS", subset(hap, INCLEVEL != HAPLEVEL))
```

The First graph is the original scatter plot, while the second is only the outliers.

Now it's time for me to figure more out about these outliers.

### Outliers

I first define a specific subset for the outliers.

Then I ran the same graphs I ran for the original set.

```{r Outlier Removal}
hap.outliers = subset(hap, INCLEVEL != HAPLEVEL)
summary(hap.outliers)

scatPlot("HAPPINESS", "INCOME", hap.outliers)
scatPlot("HAPPINESS", "AGE", hap.outliers)

plotDistr <- function (val, frm){
  table <- table(frm[,val])
  barplot(table, main=paste(val, " Distribution"), xlab="Count", col = rainbow(8), las = 2)
}

skips <- c("IDN", "INCOME", "HAPPINESS", "GEN", "AGE")

for(vart in vars){
  if(vart %in% skips){
  } else {
    plotDistr(vart, hap.outliers)
  }
}
```

And the clear similarity between all the outliers is that they're Immigrants.

And after plotting a subset of the original set, it shows that without immigrants the original trend is non existent.

Now its time to go and create the same graphs with the new immigrant subset.

### Immigrants

Creating new Immigrant subset and creating graphs

```{r Immigrants}
imgrts <- subset(hap, IMMIGRANT == 1)

imgrts[,11] <- ""
colnames(imgrts)[11] <- "NORM"
imgrts$NORM <- (as.numeric(imgrts$HAPLEVEL) == as.numeric(imgrts$INCLEVEL))

plotDistr("NORM", imgrts)

```

Most Immigrants do not fit the "Social Norm", only about 12.6% fit the normal Income/Happiness Levels.

In an attempt to discover more patterns that might show why immigrants don't fit the social norm, I re-graphed all the distributions for the Immigrants.

```{r Imgrt Distribution}

vars = colnames(imgrts)

skips <- c("IDN", "AGE", "IMMIGRANT", "INCOME", "HAPPINESS", "NORM")

for(vart in vars){
  if (vart %in% skips){
    
  } else {
    plotDistr(vart, imgrts)
  }
}
```

No patterns emerged here, so it would appear that the implanted patterns end here.

## Notes and Comments

This attempt went much more smoothly. I've begun to get a better grasp on how to approach analyzing a data set, and have made use of functions and loops to increase my plot speed.

I felt I over thought the last analysis, so I went with a more direct approach this time, and it worked out.

But I think that because this is a synthetic data set, it was easier to over think the analysis. So I look forward to dealing with more random/real world data sets in the future.

Our next project won't involve a synthetic data set, so I look forward to the new challenges that the next step in analysis presents.

---

[Rutgers Data 101 Blog](http://data101.cs.rutgers.edu/?q=blog)

[My Data Exploration Homepage](home.html)

[Rutgers Data 101 Homepage](http://data101.cs.rutgers.edu/)
