---
title: "Census Investigation"
author: "David Parsons"
date: "March 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# # library(ggplot2)
# install.packages("devtools")
devtools::install_github("devanshagr/PermutationTestSecond")
options(scipen=999)
```

---

[Rutgers Data 101 Blog](http://data101.cs.rutgers.edu/?q=blog)

[My Data Exploration Homepage](home.html)

[Rutgers Data 101 Homepage](http://data101.cs.rutgers.edu/)

---

## Introduction

This test is in response to a course assignment for my CS 142 class. We we're provided with a synthetic data set that had artificial patterns created with in it for us to find in a [previous assignment](Census_Investigation.html).

No we have been tasked with supporting or rejecting hypotheses that were found about this using permutation testing.

## The Data

Data Provided by Professor Imenlinski as part of a CS 142 Assignment. And is exactly the same as the data used [previously](Census_Investigation.html) except for one added category shown later.

```{r The Data}
census <- read.csv("CENSUSNEW.csv")
summary(census)
```
    
I did some initial set creation so I could more easily iterate over the categories of the data frame. This came in handy later when created my own functions and used for-each loops to create lots of plots with little repetition on my part.

```{R Initial Set Up}

census[, 11] <- ""
colnames(census)[11] <- "isDoc"
census$isDoc <- (census$EDUCATION == "Doctorate")

# Creates Dictionary of Numerics
nums <- c("AGE", "YEARS", "CAPITALNET")
# Creates Dictionary of Elements
elems <- levels(census$ELEMENTS)
# Non Element cats
cats <- c("AGE", "YEARS", "CAPITALGAINS", "CAPITALLOSS", "CAPITALNET")
```

Here I create an "isDoc" category, that defines whether or not a respondent holds a doctorate or not.

```{R Capital Manipulations}
# Creates Net Capital Variable
census$CAPITALNET = census$CAPITALGAINS - census$CAPITALLOSS
# Summary with CAPITALNET
summary(census)
```
### Testing

The following runs permutation tests on the following hypotheses:
    H1
      Doctorate holders are older than those at all other education levels.
      What types of people have doctorates?
    H2
      Doctorate holders have higher Net Capital than those at all other education levels.
      Is getting a doctorate worth it financially?

```{R Testing}
# Permuatation testing for Age and Net Capital
print(PermutationTestSecond::Permutation(census, "isDoc", "AGE", 50000, FALSE, TRUE))
print(PermutationTestSecond::Permutation(census, "isDoc", "CAPITALNET", 50000, FALSE, TRUE))
# Flipped Parameters Sanity Test/Check
# print("Flipped")
# print(PermutationTestSecond::Permutation(census, "isDoc", "AGE", 50000, TRUE, FALSE))
# print(PermutationTestSecond::Permutation(census, "isDoc", "CAPITALNET", 50000, TRUE, FALSE))
```

### H1 - Age Test

With a p value of 0, it is very easy to accept that Doctorate holders are in fact older than those at other education levels. This sort of makes sense intuitively, since doctorates require more time and experience to complete.

### H2 - Capital Gains Test

On the other hand with a p value of ~0.38, much greater than the 0.05 alpha, it's much harder to accept that having a doctorate has a direct effect on an increase on Net Capital.

## Notes and Comments

While I found this exercise in permutation a bit trivial (because the data set is synthetic), working with p-values has given me a better appreciation for data investigation. Despite what patterns may appear to exist, there's always the element of chance, and its important for people to understand & recognize this so they aren't easily swayed by quick facts and figures.

I looked forward to incorporating this and other types of testing into my data investigations in the future.

---

[Rutgers Data 101 Blog](http://data101.cs.rutgers.edu/?q=blog)

[My Data Exploration Homepage](home.html)

[Rutgers Data 101 Homepage](http://data101.cs.rutgers.edu/)