---
title: "Prediction Challenge 3"
author: "David Parsons"
date: "April 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("rpart")
# install.packages("rpart.plot")
# install.packages("e1071")
install.packages("devtools")
devtools::install_github("devanshagr/CrossValidation")
library(rpart)
library(rpart.plot)
options(scipen=999)
```

---

[Rutgers Data 101 Blog](http://data101.cs.rutgers.edu/?q=blog)

[My Data Exploration Homepage](home.html)

[Rutgers Data 101 Homepage](http://data101.cs.rutgers.edu/)

---

## Introduction

This investigation is in response to a course assignment for my CS 142 class. We we're provided with a synthetic data set that had artificial patterns created with in it for us to find.

## The Data

Data Provided by Professor Imenlinski as part of a CS 142 Assignment.

```{r The Data}
input <- read.csv("Marriage_training.csv")
test <- read.csv("Marriage_testing_students1.csv")

print(summary(train))
```

```{r rPart & Cross Validation}

# Partition Data
train <- input[1:349,]
val <- input[350:750,]

# Creating Prediction Tree
tree <- rpart(STATUS ~ (Joint.Income+(GroomPersonality==BridePersonality)+(GroomAge<BrideAge)+(GroomAge==BrideAge)), control = rpart.control(minsplit = 5), data = train)

summary(tree)
rpart.plot(tree)

test.tree <- predict(tree, newdata=val, type="class")
mean(val$STATUS == test.tree)

# Validation Testing
CrossValidation::cross_validate(val, tree, 1000, 0.5)

# Prediction with test data
test_predict <- predict(tree, newdata = test, type = "class")
summary(test_predict)


```

### Output

```{r Output}

# To File
submission<-read.csv("Marriage_sample.csv")
submission$STATUS<-test_predict
write.csv(submission, 'submission.csv', row.names= FALSE)

```

## Notes and Comments

This attempt went much more smoothly. I've begun to get a better grasp on how to approach analyzing a data set, and have made use of functions and loops to increase my plot speed.

I felt I over thought the last analysis, so I went with a more direct approach this time, and it worked out.

But I think that because this is a synthetic data set, it was easier to over think the analysis. So I look forward to dealing with more random/real world data sets in the future.

Our next project won't involve a synthetic data set, so I look forward to the new challenges that the next step in analysis presents.

---

[Rutgers Data 101 Blog](http://data101.cs.rutgers.edu/?q=blog)

[My Data Exploration Homepage](home.html)

[Rutgers Data 101 Homepage](http://data101.cs.rutgers.edu/)