---
title: "Prediction Challenge 5"
author: "David Parsons"
date: "April 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("rpart")
# install.packages("rpart.plot")
# install.packages("e1071")
# install.packages("devtools")
# devtools::install_github("devanshagr/CrossValidation")
library(rpart)
library(rpart.plot)
options(scipen=999)
```

---

[Rutgers Data 101 Blog](http://data101.cs.rutgers.edu/?q=blog)

[My Data Exploration Homepage](home.html)

[Rutgers Data 101 Homepage](http://data101.cs.rutgers.edu/)

---

## Introduction

This investigation is in response to a course assignment for my CS 142 class. We we're provided with a synthetic data set that had artificial patterns created with in it for us to find.

## The Data

Data Provided by Professor Imenlinski as part of a CS 142 Assignment.

```{r The Data}
input <- read.csv("P5_train1.csv")
# input <- subset(input, input$Source != "Directlink")
# test <- read.csv("Marriage_testing_students1.csv")
partitionData <- function(){
  samp.total <- sample(nrow(input), 4488, replace = FALSE)
  samp.train <- sample(nrow(input), 1795, replace = FALSE)
  samp.test <- sample(nrow(input[setdiff(samp.total, samp.train),]), 2244, replace = FALSE)
  
  train <- input[samp.train,] # random 40% of data
  test <- input[samp.test,] # random 50% of data
}
print(summary(input))
```

```{r rPart & Cross Validation}

MSE <- 100
for(j in c(3:17)){
  MSE <- 100
    
  tree.lm <- lm(Revenue ~ ., data = input)
  for(i in c(1:5000)){
    samp.total <- sample(nrow(input), 4950, replace = FALSE)
    samp.train <- sample(nrow(input), (4950 * (j/20)), replace = FALSE)
    samp.test <- sample(nrow(input[setdiff(samp.total, samp.train),]), (4950 * ((18 - j + 1)/20)), replace = FALSE)
    
    train <- input[samp.train,] # random 40% of data
    test <- input[samp.test,] # random 50% of data
    
    tmpTree.lm <- lm(Revenue ~ ., data = train)
    
    # summary(tree)
    # rpart.plot(tree)
    
    outVec <- predict(tmpTree.lm, newdata=test)
    # mean(test$Revenue == test.tree)
    
    tmpMSE <- mean((outVec - test$Revenue)^2)
    
    if(tmpMSE < MSE){
      MSE <- tmpMSE
      tree.lm <- tmpTree.lm
    }
    
  }
  print(paste("Train:", (j/20), "Test:", ((18 - j)/20), " -> ", MSE))
  # plot(tree.lm)
}

# Prediction with test data
# test_predict <- predict(tree, newdata = test, type = "class")
# summary(test_predict)


```

### Output

```{r Output}

MSE <- 100

tree.lm <- lm(Revenue ~ ., data = input)
for(i in c(1:100000)){
  samp.total <- sample(nrow(input), 4950, replace = FALSE)
  samp.train <- sample(nrow(input), (4950 * 0.65), replace = FALSE)
  samp.test <- sample(nrow(input[setdiff(samp.total, samp.train),]), (4950 * 0.25), replace = FALSE)
  
  train <- input[samp.train,] # random 40% of data
  test <- input[samp.test,] # random 50% of data
  
  tmpTree.lm <- lm(Revenue ~ ., data = train)
  
  # summary(tree)
  # rpart.plot(tree)
  
  outVec <- predict(tmpTree.lm, newdata=test)
  outVec[outVec < 0] <- 0
  # mean(test$Revenue == test.tree)
  
  tmpMSE <- mean((outVec - test$Revenue)^2)
  
  if(tmpMSE < MSE){
    MSE <- tmpMSE
    tree.lm <- tmpTree.lm
  }

}
print(MSE)

# Predicting on Real Data
data <- read.csv("P5_test_students.csv")
finalVec <- predict(tree.lm, newdata = data)
finalVec[finalVec < 0] <- 0

# To File
submission<-read.csv("P5_sample.csv")
submission$Revenue <- finalVec
write.csv(submission, 'submission.csv', row.names= FALSE)

```

## Notes and Comments

This attempt went much more smoothly. I've begun to get a better grasp on how to approach analyzing a data set, and have made use of functions and loops to increase my plot speed.

I felt I over thought the last analysis, so I went with a more direct approach this time, and it worked out.

But I think that because this is a synthetic data set, it was easier to over think the analysis. So I look forward to dealing with more random/real world data sets in the future.

Our next project won't involve a synthetic data set, so I look forward to the new challenges that the next step in analysis presents.

---

[Rutgers Data 101 Blog](http://data101.cs.rutgers.edu/?q=blog)

[My Data Exploration Homepage](home.html)

[Rutgers Data 101 Homepage](http://data101.cs.rutgers.edu/)